---
title: "categorical"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{categorical}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library("colorpath")
library("purrr")
library("glue")
library("ggplot2")
library("tibble")
```

The goal of this vignette is to show a way of building a categorical color-palette:

  - eight colors
  - paired (four sets of two)
  - designed to be useful under color-vision deficiency
  
We will use the four surfaces introduced in the [surfaces](surfaces.html) vignette.

```{r surfaces}
# TODO: find a more-robust way to keep this in sync.
colors <- list(
  blues = c("#42B4E6", "#0087CD"),
  oranges = c("#E47F00", "#702407"),
  greens = c("#007626", "#70E07C"),
  purples = c("#3D1152", "#C530FF")
)

surfaces <- purrr::map(colors, pth_new_surface, transformer = pth_to_jzazbz100)
```

We can look at how the outer edge of these surfaces "collapse" under color-vision deficiency:

```{r mat_surfaces}
mat_surfaces <- 
  surfaces %>%
  map(pth_mat_max_chroma) %>%
  reduce(rbind)
```

Next, we can plot these colors in polar coordinates, using `pth_plot_polar()`:

```{r plot-mat_surfaces}
pth_plot_polar(mat_surfaces, cvd = pth_cvd_grid_full()) +
  facet_grid(
    rows = vars(condition), 
    cols = vars(severity),
    labeller = labeller(severity = label_both)
  ) +
  ggtitle(
    label = "Outer (chroma) edge of surfaces, simulating color-vision deficiency",
    subtitle = glue("Color space: {pth_colorspace_name(mat_surfaces)}")
  )
```

I am going to propose a design frameowrk, based on deuteranopia, because:

- deuteranopia (green deficiency) is the predominant form of color-vision deficiency.
- it manifests itself similarly to protanopia.
- tritanopia still seems to leave some variation in hue.

In short, I propose to choose:

- two blues and two purples in such a way that these colors do not collide under deuteranopia.
- two greens and two oranges, also not to collide under deuteranopia.

It is clear that under full deuteranopia, for blues and purples, there is no variation in hue. 
Therefore, we will have to make things different using luminance and chroma.

```{r blues-purples-01}
blues_purples_first <-
  tribble(
        ~name, ~luminance, ~saturation,
      "blues",       80.0,        1.00, 
    "purples",       65.0,        1.00, 
    "purples",       45.0,        1.00, 
      "blues",       30.0,        1.00
  )
```

```{r function-cat}
# look at making this function a part of the package...
get_color_cat <- function(surface, luminance, saturation = 1) {
  
  chroma <- surface$fn_max_chroma(luminance) * saturation
  hue <- surface$fn_hue(luminance)
  
  mat_polar <- matrix(c(luminance, chroma, hue), ncol = 3)
  mat_cartesian <- colorpath::pth_to_cartesian(mat_polar)
  
  color <- colorpath::pth_creator(surface$colors)(mat_cartesian)
  
  color
}

get_color_cat_surfaces <- function(name, luminance, saturation = 1) {
  get_color_cat(surfaces[[name]], luminance, saturation)
}
```

```{r}
blues_purples_mat <- 
  pmap(blues_purples_first, get_color_cat_surfaces) %>%
  reduce(rbind)
```


```{r}
blues_purples_cvd <-
  pth_data_cvd(
    blues_purples_mat, 
    cvd = pth_cvd_grid(
      condition = c("deutan", "protan"), 
      severity = seq(0, 1, 0.1)
    )
  )
```

```{r}
cat_plot <- function(mat) {
 
  keep <- function(x) {
    keep <- x$severity == max(x$severity)
    x[keep, ]
  }

  ggplot2::ggplot(mat, aes(x = chroma, y = luminance)) +
    ggplot2::geom_point(aes(color = hex, shape = factor(hex_original)), size = 4) + 
    ggplot2::geom_point(data = keep, shape = 3, color = "black", size = 4) +
    ggplot2::geom_point(data = keep, shape = 4, color = "white", size = 4) +
    ggplot2::facet_wrap(facets = "condition") +
    ggplot2::scale_color_identity() +
    ggplot2::scale_shape_manual(values = c(15, 16, 17, 18), guide = NULL) + 
    ggplot2::coord_fixed(ratio = 1, xlim = c(0, NA), ylim = c(0, 100)) +
    ggplot2::theme_bw() + 
    ggplot2::labs(
      title = "Colors under color-vision deficency of varying severity",
      subtitle = "Cross-marks indicate maximum severity"
    ) 
}

cat_plot(blues_purples_cvd)
```

Here, we can see the measures we can take to keep these four colors more separated. 
We can see that, according to the color-vision deficiency models, the colors with the purple hue drop in luminance (and increase in chroma) as the severity increases.

With this in mind, we can choose the "original" colors to improve the separation at maximum severity.

```{r blues-purples-02}
blues_purples_second <-
  tribble(
        ~name, ~luminance, ~saturation,
      "blues",       75.0,        0.75, 
    "purples",       65.0,        0.77, 
    "purples",       35.0,        1.00, 
      "blues",       45.0,        1.00
  )
```

```{r}
blues_purples_mat_second <- 
  pmap(blues_purples_second, get_color_cat_surfaces) %>%
  reduce(rbind)
```


```{r}
blues_purples_cvd_second <-
  pth_data_cvd(
    blues_purples_mat_second, 
    cvd = pth_cvd_grid(
      condition = c("deutan", "protan"), 
      severity = seq(0, 1, 0.1)
    )
  )
```

```{r}
cat_plot(blues_purples_cvd_second)
```

We can do the same thing with greens and oranges:

```{r greens_oranges}
greens_oranges <-
  tribble(
        ~name, ~luminance, ~saturation,
     "greens",       65.0,        0.65, 
    "oranges",       65.0,        0.90, 
    "oranges",       30.0,        1.00, 
     "greens",       40.0,        0.80
  )
```

```{r}
greens_oranges_mat <- 
  pmap(greens_oranges, get_color_cat_surfaces) %>%
  reduce(rbind)
```


```{r}
greens_oranges_cvd <-
  pth_data_cvd(
    greens_oranges_mat, 
    cvd = pth_cvd_grid(
      condition = c("deutan", "protan"), 
      severity = seq(0, 1, 0.1)
    )
  )
```

```{r}
cat_plot(greens_oranges_cvd)
```

Let's get these colors together:

```{r}
categorical_hex <- 
  rbind(blues_purples_mat_second, greens_oranges_mat) %>%
  `[`(c(4, 6, 3, 5, 1, 7, 2, 8), ) %>%
  pth_to_hex

colorspace::swatchplot(categorical_hex, cvd = TRUE)
```

We need a function that, given a set of colors (hex or mat), returns a data frame with variables:

- hex_original_a
- hex_original_b
- cvd
- severity
- hex_a
- hex_b
- distance

```r
# suggests
pth_distance_color_euclid(color_a, color_b, cvd, transformer, non_luminance_weight, ...)
pth_distance_color_metric(color_a, color_b, cvd, method, ...)

pth_distance_cat_euclid(color, cvd, transformer, non_luminance_weight, ...)
pth_distance_cat_metric(color, cvd, method, ...)

pth_distance_qnt_euclid(palette, n_step, cvd, transformer, non_luminance_weight, ...)
pth_distance_qnt_metric(palette, n_step, cvd, method, ...)
```
