---
title: "Using Color Surfaces to Incorporate an Identity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{surfaces}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r opts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library("colorpath")
```

One goal of this package is to help you design sets of categorical and quantitative palettes that are consistent amongst themselves.

We propose a mechanism for this, what this package calls a *surface*.

## Single-hue surface

At it's simplest, a surface is the set of all colors (in a given color space) with a given hue. 
Let's consider a dark-blue hue in the Jzazbz color-space:

```{r sfc_blue}
sfc_blue <- pth_new_surface("#0000FF", transformer = pth_to_jzazbz100)
```

It takes a few seconds for the surface to be computed; it is calculating values for maximum chroma. 
Here's a plot of the result:

```{r sfc_blue-plot}
pth_plot_surface(sfc_blue)
```

This surface is a slice of the RGB gamut for the Jzazbz color-space (scaled here to 100) where the hue is fixed to the hue of `#0000FF`.

We can make a similar surface for cyan (`#00FFFF`):

```{r sfc_cyan}
sfc_cyan <- pth_new_surface("#00FFFF", transformer = pth_to_jzazbz100)
pth_plot_surface(sfc_cyan)
```

We notice a couple of things here:

- the maximum chroma for blue is at a low luminance.
- the maximum chroma for cyan is at a high luminance.

## Multi-hue surfaces

We can create a surface that captures both these maxima:

```{r sfc_cyan_blue}
sfc_cyan_blue <- pth_new_surface(c("#00FFFF", "#0000FF"), transformer = pth_to_jzazbz100)
pth_plot_surface(sfc_cyan_blue)
```

This shows another aspect of surfaces: using the given color-space, hue is a linear function of luminance. 
This was also the case for the single-hue surfaces.

Also note that you can define surfaces using hex-codes.
This could be useful if you are designing a set of palettes for an organization, you can design surfaces using the hex-codes of existing brand-colors.

When you create a surface using hex-codes, you have to specify the color-space to be used; provide a `transformer` function such as `pth_to_jzazbz100`, without the parentheses.

There are couple other arguments to `pth_new_surface()`:

- use `n_step` to specify the number of steps, in luminance, to use in calculating the maximum chroma. 
  The max-chroma is calculated *once*, as the surface is being created; a much quicker approximation-function is used thereafter.
  
- use `route` to specify if the variation of hue - to we take the `"short"` (default) or `"long"` route around the unit circle?
  The `"long"` value may be useful if you are designing a surface to build a viridis-like or inferno-like palette.

You can see that the surface `sfc_blue_cyan` starts to look "purple" for values of luminance below around 40. 
We can make the surface more "blue" at the low end by choosing a different dark blue, in this case `"#000077"`:
  
```{r sfc_cyan_blue_clipped}
sfc_cyan_blue_clipped <- pth_new_surface(c("#00FFFF", "#000077"), transformer = pth_to_jzazbz100)
pth_plot_surface(sfc_cyan_blue_clipped)
```  

This surface looks more "blue" top-to-bottom.

## Incorporating an Itendtity
  

